Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-12-08 src/tex/ms.tex
Updates from Overleaf

The two-infall scenario reproduces the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lian_quantifying_2022,lehmann_probing_2024}. 
A number of GCE studies have reproduced the $\alpha$-bimdodality using a combination of radial migration and smooth, inside-out galaxy growth \citep[e.g.,][]{schonrich_chemical_2009,kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history (although the models of \citealt{johnson_stellar_2021} and \citealt{dubay_galactic_2024} failed to produce an $\alpha$-bimodality).
Figure \ref{fig:smooth-vs-twoinfall} contrasts the chemical evolution track for the Solar neighborhood predicted by a two-infall model against tracks from the smooth SFH model of \citet{johnson_stellar_2021} at three different radii. 
If the effect of radial migration is significant, then the local $\alpha$-bimodality could be explained without dramatic changes to the local gas abundance over time.

@@

\todo{Integrate into previous para:} \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation (AMR), which could be interpreted as evidence supporting the two-infall scenario, but \citet{plotnikova_chemical_2024} argued that such patterns can arise due to selection effects.
\todo{Also: multiple sequences in Anders et al. (2023), Xiang \& Rix (2022), Sahlholdt et al. (2022).}

The availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2025} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}. We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we and summarize our conclusions in Section \ref{sec:conclusions}.

@@

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=\onecolumn]{\figpath{gas_abundance_evolution.pdf}}
    \caption{The ISM abundance evolution at $R_{\rm gal}=8\kpc$ of three multi-zone models at different yield scales (see Table \ref{tab:yields}), each of which feature a major dilution event 9 Gyr ago.
    % The dark gray (light gray) shaded region plots the 68\textsuperscript{th} (95\textsuperscript{th}) percentile of APOGEE abundances in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $|z|<0.5\kpc$) with NN ages from \citet{leung_variational_2023}. The black line (black points) plots the median (mode) of the abundance data in 1 Gyr-wide age bins, and the gray error bars along the bottom of each panel indicate the median age and abundance errors as a function of age. 
    \added{The 2-D histograms show the APOGEE stellar age--abundance distributions in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $|z|<0.5\kpc$), adopting the \citet{leung_variational_2023} ages. The black curve (points) plots the rolling median (binned mode) of the abundance data as a function of age, and the gray error bars along the bottom of each panel indicate the median age and abundance errors as a function of age.}
    The left-hand marginal panels show the predicted (blue, pink, and green) and observed (gray) stellar abundance distributions \added{in the same region}, which are boxcar-smoothed with a width of 0.05 dex for visual clarity.}
    \label{fig:yield-outflow}
\end{figure}

Figure \ref{fig:yield-outflow} shows the \added{APOGEE age--abundance distributions at the Solar annulus, as well as the binned mode abundance and rolling median as a function of age}, using the NN ages from \citet{leung_variational_2023}. As shown by \citet{johnson_milky_2025}, the peak of the MDF is less susceptible to modification by radial migration, making it a more reliable proxy for ISM chemistry at a given lookback time than the mean or median. \added{While the age--metallicity distribution is clumpy in places (see Section \ref{sec:amr}), we observe that ${\rm mode}(\mathOH)\gtrsim0$ for the past 10 Gyr.}

@@

Motivated by the results of the previous sections, we attempt to construct a model that solves all of the issues that have been outlined thus far. Our ``best attempt'' model {\tt yZ2-best} uses the \yZ{2} yield set to flatten the local age--metallicity relation (Figure \ref{fig:yield-outflow}), pre-enriched infall at the level of $\mathXH_{\rm CGM}=-0.7$ to reduce the dilution at $t_{\rm max}$ (Figure \ref{fig:abundance-evolution-params}), slightly stronger outflows with $\eta_\odot=1.8$ to maintain the local equilibrium at Solar metallicity, moderately stronger radial migration with $\sigma_{\rm RM8}=3.6\kpc$ to widen the local metallicity dispersion (Figure \ref{fig:abundance-evolution-params}), and a greater local disk ratio $f_\Sigma(R_\odot)=0.25$ to reduce the width of the low-$\alpha$ distribution and beef up the high-$\alpha$ sequence (Figure \ref{fig:ofe-distribution-parameters}). Our choices for $\mathXH_{\rm CGM}$, $\sigma_{\rm RM8}$, and $f_\Sigma(R_\odot)$ are more moderate than the illustrative examples in the previous section. Our focus is on qualitative rather than quantitative agreement with the data, and thus we do not attempt to find the optimal set of parameters.

Figure \ref{fig:ofe-feh-best} shows the stellar [O/Fe]--[Fe/H] distributions in different ranges of $R_{\rm gal}$ and $|z|$ color-coded by age as predicted by the {\tt yZ2-best} model. This model is generally successful at reproducing the observed abundance distributions, especially in the inner Galaxy and close to the midplane. However, the predicted high-$\alpha$ sequence has a significant presence even in the outer Galaxy---likely a consequence of the stronger migration prescription and higher thick-to-thin disk ratio. In general, the predicted distributions do not align with the data as well at large midplane distances ($1\leq|z|<2\kpc$), but this may reflect inaccuracies in our prescription for vertical heating (see Section \ref{sec:migration}).

@@

\added{A few recent studies have reported the discovery of multiple disconnected sequences in the local age--metallicity relation \citep[e.g.,][]{sahlholdt_characterizing_2022,xiang_time-resolved_2022,anders_spectroscopic_2023}. \citet{nissen_high-precision_2020} argued that these sequences provide evidence for the two-infall scenario, although \citet{plotnikova_chemical_2024} showed that multiple sequences can arise due to selection effects, and \citet{chen_recent_2025} showed that radial migration can also lead to such patterns. There are hints of multiple sequences in Figure \ref{fig:yield-outflow}, where a break in the distribution at ages of $\sim5\Gyr$ coincides with a $\sim0.2\dex$ drop in [O/H] and [Fe/H]. The sequences are less pronounced in the \citet{leung_variational_2023} catalog than reported by \citet{anders_spectroscopic_2023}, likely related to a systematic offset between the two catalogs around the same age \citep[see Figure 6 of][]{anders_spectroscopic_2023}. On the other hand, the [C/N] age-metallicity relation of \citet{roberts_cn_2025} is even flatter and does not appear to have multiple sequences. Even if these sequences are real, we find it difficult to explain them using the two-infall scenario\textemdash the observed dilution is much smaller than predicted by any of our models, and the timing of the split between the sequences is too late for the birth of the thin disk, although it is interestingly close to the first pericentric passage of the Sagittarius dwarf galaxy \citep{ruiz-lara_recurrent_2020}.}
\todo{Mention APOGEE selection function?}
% Anders+ use APOGEE clump stars. Their split is 5-6 Gyr ago
% Xiang & Rix use Gaia subgiants. Their split is 8 Gyr ago
% Sahlholdt+ use GALAH turnoff stars. Their split is 7-9 Gyr ago.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{\figpath{compare_age_catalogs.pdf}}
    \caption{Comparison between NN- and [C/N]-derived ages for our sample.}
    \label{fig:compare-age-catalogs}
\end{figure*}

>>>>

The two-infall scenario reproduces the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. 
However, 
% the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and 
it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{sellwood_radial_2002,schonrich_chemical_2009,frankel_measuring_2018,lian_quantifying_2022,lehmann_probing_2024}. 
A number of GCE studies have reproduced the $\alpha$-bimdodality using a combination of radial migration and smooth, inside-out galaxy growth (e.g., \citealt{schonrich_chemical_2009,kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}; \added{but see also \citealt{johnson_stellar_2021,dubay_galactic_2024}}). In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history.
% (although the models of \citealt{johnson_stellar_2021} and \citealt{dubay_galactic_2024} failed to produce an $\alpha$-bimodality).
Figure \ref{fig:smooth-vs-twoinfall} contrasts the chemical evolution track for the Solar neighborhood predicted by a two-infall model against tracks from the smooth SFH model of \citet{johnson_stellar_2021} at three different radii. 
If the effect of radial migration is significant, then the local $\alpha$-bimodality could be explained without dramatic changes to the local gas abundance over time.

@@

The availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2025} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}. We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we summarize our conclusions in Section \ref{sec:conclusions}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{\figpath{compare_age_catalogs.pdf}}
    \caption{\added{Comparison between NN \citep{leung_variational_2023} and [C/N] \citep{roberts_cn_2025} ages for our sample. {\it Left:} Comparison of age estimates from both methods for all APOGEE stars. {\it Right:} The local ($7\leq R_{\rm gal}<9\kpc$, $|z|<0.5\kpc$) age--metallicity relation for each age estimation method.}}
    \label{fig:compare-age-catalogs}
\end{figure*}

@@

\added{Figure \ref{fig:compare-age-catalogs} compares the NN ages from \citet{leung_variational_2023} and the [C/N]-derived ages from \citet{roberts_cn_2025} for our sample. Below an age of $\sim10\Gyr$, the two methods trace each other closely, although there is a small systematic offset for $\lesssim4\Gyr$ old stars. The Solar neighborhood age--metallicity relation (AMR) of \citet{leung_variational_2023} features substructure not seen in the \citet{roberts_cn_2025} AMR, possibly washed out due to the larger uncertainty in the latter sample. We compare these AMRs to others in the literature in Section \ref{sec:amr}. Overall, the AMRs derived from both methods are remarkably flat out to $\sim8-9\Gyr$, with variations of $\lesssim0.1\dex$ on the order of $1\Gyr$.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\onecolumn]{\figpath{gas_abundance_evolution.pdf}}
    \caption{The ISM abundance evolution at $R_{\rm gal}=8\kpc$ of three multi-zone models at different yield scales (see Table \ref{tab:yields}), each of which feature a major dilution event 9 Gyr ago.
    % The dark gray (light gray) shaded region plots the 68\textsuperscript{th} (95\textsuperscript{th}) percentile of APOGEE abundances in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $|z|<0.5\kpc$) with NN ages from \citet{leung_variational_2023}. The black line (black points) plots the median (mode) of the abundance data in 1 Gyr-wide age bins, and the gray error bars along the bottom of each panel indicate the median age and abundance errors as a function of age. 
    \added{The 2-D histograms show the APOGEE stellar age--abundance distributions in the Solar neighborhood ($7\leq R_{\rm gal}<9\kpc$, $|z|<0.5\kpc$), adopting the \citet{leung_variational_2023} ages. The black curve (points) plots the rolling median (binned mode) of the abundance data as a function of age, and the gray error bars along the bottom of each panel indicate the median age and abundance errors as a function of age.}
    The left-hand marginal panels show the predicted (blue, pink, and green) and observed (gray) stellar abundance distributions, which are boxcar-smoothed with a width of 0.05 dex for visual clarity.}
    \label{fig:yield-outflow}
\end{figure}

Figure \ref{fig:yield-outflow} shows the \added{APOGEE age--abundance distributions in the Solar annulus, as well as the binned mode and rolling median abundance as a function of age}, using the NN ages from \citet{leung_variational_2023}. \added{While the age--abundance distributions are clumpy in places (see Section \ref{sec:amr}), we caution that the APOGEE selection function can affect the stellar age distribution \citep[see][]{imig_galactic_2025}.} As shown by \citet{johnson_milky_2025}, the peak of the MDF is less susceptible to modification by radial migration, making it a more reliable proxy for ISM chemistry at a given lookback time than the mean or median. \added{Strikingly, we observe that ${\rm mode}(\mathOH)\gtrsim0$ over the past 10 Gyr.}

@@

Motivated by the results of the previous sections, we attempt to construct a model that solves all of the issues that have been outlined thus far. Our ``best attempt'' model {\tt yZ2-best} uses the \yZ{2} yield set to flatten the local age--metallicity relation (Figure \ref{fig:yield-outflow}), \added{$t_{\rm max}=3.2\Gyr$ to better match the local [O/Fe]--[Fe/H] distribution (Section \ref{sec:sfh}),} pre-enriched infall at the level of $\mathXH_{\rm CGM}=-0.7$ to reduce the dilution at $t_{\rm max}$ (Figure \ref{fig:abundance-evolution-params}), slightly stronger outflows with $\eta_\odot=1.8$ to maintain the local equilibrium at Solar metallicity, moderately stronger radial migration with $\sigma_{\rm RM8}=3.6\kpc$ to widen the local metallicity dispersion (Figure \ref{fig:abundance-evolution-params}), and a greater local disk ratio $f_\Sigma(R_\odot)=0.25$ to reduce the width of the low-$\alpha$ distribution and beef up the high-$\alpha$ sequence (Figure \ref{fig:ofe-distribution-parameters}). Our choices for $\mathXH_{\rm CGM}$, $\sigma_{\rm RM8}$, and $f_\Sigma(R_\odot)$ are more moderate than the illustrative examples in the previous section. Our focus is on qualitative rather than quantitative agreement with the data, and thus we do not attempt to find the optimal set of parameters.

Figure \ref{fig:ofe-feh-best} shows the stellar [O/Fe]--[Fe/H] distributions in different ranges of $R_{\rm gal}$ and $|z|$ color-coded by age as predicted by the {\tt yZ2-best} model. This model is generally successful at reproducing the observed abundance distributions, especially in the inner Galaxy and close to the midplane. However, the predicted high-$\alpha$ sequence has a significant presence even in the outer Galaxy---likely a consequence of the stronger migration prescription and higher thick-to-thin disk ratio \added{than in the fiducial model}. In general, the predicted distributions do not align with the data as well at large midplane distances ($1\leq|z|<2\kpc$), but this may reflect inaccuracies in our prescription for vertical heating (see Section \ref{sec:migration}).

@@

\added{A few recent studies have reported the discovery of multiple disconnected sequences in the local age--metallicity relation \citep[e.g.,][]{sahlholdt_characterizing_2022,xiang_time-resolved_2022,anders_spectroscopic_2023}. \citet{nissen_high-precision_2020} argued that these sequences provide evidence for the two-infall scenario, although \citet{plotnikova_chemical_2024} showed that multiple sequences can arise due to selection effects, and \citet{chen_recent_2025} showed that radial migration can also lead to such patterns. There are hints of multiple sequences in Figure \ref{fig:yield-outflow}, where a break in the distribution at ages of $\sim5\Gyr$ coincides with a $\sim0.2\dex$ drop in [O/H] and [Fe/H]. The sequences are less pronounced in the \citet{leung_variational_2023} catalog than reported by \citet{anders_spectroscopic_2023}, likely related to a systematic offset between the two catalogs around the same age \citep[see Figure 6 of][]{anders_spectroscopic_2023}. On the other hand, the [C/N] age--metallicity relation of \citet{roberts_cn_2025} is even flatter and does not appear to have multiple sequences (Figure \ref{fig:compare-age-catalogs}). Even if these sequences are real, we find it difficult to explain them using the two-infall scenario\textemdash the observed dilution is much smaller than predicted by any of our models, and the timing of the split between the sequences is too late for the birth of the thin disk, although it is interestingly close to the first pericentric passage of the Sagittarius dwarf galaxy \citep{ruiz-lara_recurrent_2020}.}
\todo{Mention APOGEE selection function?}
% Anders+ use APOGEE clump stars. Their split is 5-6 Gyr ago
% Xiang & Rix use Gaia subgiants. Their split is 8 Gyr ago
% Sahlholdt+ use GALAH turnoff stars. Their split is 7-9 Gyr ago.

\todo{Even if the Anders et al. (2023) sequences are right, it still doesn't look like two-infall. The Anders AMR reaches Solar-ish metallicity 8 Gyr ago, then drops to $\mathFeH\approx-0.2$ by 4 Gyr ago. A 0.2 dex dilution over 4 Gyr is {\it far} weaker and more gradual than predicted by any of our two-infall models.}
